---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r, include = FALSE}
# library(rogme)
devtools::load_all()
library(ggplot2)
library(tibble)
```

# garstats v0.0.0.9

<!-- badges: start -->
<!-- badges: end -->

`garstats` is an R package to apply and teach robust inferential methods. The goal is to cover basic parametric and percentile bootstrap methods for group comparisons. Plot and utility functions help illustrate bootstrap results.

## Installation

You can install the development version of `garstats` from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("GRousselet/garstats")
```

## Content of the R files

### `wilcoxmod.R`

The functions in this file were modified from the original code from Rand R. Wilcox:
http://dornsife.usc.edu/labs/rwilcox/software/
https://osf.io/xhe8u/
The main reference for the code is:
Wilcox, Rand R. 2022. Introduction to Robust Estimation and Hypothesis Testing. 5th edn.
Statistical Modeling and Decision Science. San Diego, CA: Academic Press.

Some of the main changes include: 

- do not remove missing values inside the function

- change default bootstrap samples to 2000

- add option to return two-sided or one-sided p-values and confidence intervals

- remove call set.seed() in the functions; it is better practice to set the seed outside of the functions

- use quantile() to compute quantiles -- default type=6

- export bootstrap distributions

- boostrap p-values are by default corrected for small sample sizes

### `plot.R`

Functions to plot the bootstrap distributions generated by some of the functions in `wilcoxmod.R`.

### `quantiles.R`

Functions to compute various distributional characteristics captured by combinations of quantiles, such as:

- location 

- spread

- skewness

- kurtosis

- inequality

All functions are based on the Harrell-Davis quantile estimator and are inspired by the functions in the `rquest` package.
Reference:
Prendergast, L. A., Dedduwakumara, S., & Staudte, R. G. (2024). 
rquest: An R package for hypothesis tests and confidence intervals for quantiles and summary measures based on quantiles (arXiv:2410.11093; Version 1). arXiv. https://doi.org/10.48550/arXiv.2410.11093

### `utils.R`

Utility functions including a default `ggplot2` theme and a colourmap.

# Examples of percentile bootstrap inferences

## Generate data and plot densities
```{r}
set.seed(666)
n <- 50
g1 <- rbeta(n, shape1 = 2, shape2 = 2)
g2 <- rbeta(n, shape1 = 6, shape2 = 6)

df <- tibble(x = c(g1, g2),
             Group = factor(rep(c("Group 1", "Group 2"), each = n)))

ggplot(df, aes(x = x, colour = Group, linetype = Group)) + theme_minimal() +
  geom_line(stat = "density", linewidth = 1.5) + 
  scale_colour_manual(values = c("black", "grey")) +
  scale_linetype_manual(values = c("longdash", "solid"))
```

### Plot CDFs
```{r}
plot.ecdf(g1, g2)
```

## One-sample inference 

### Two-sided
```{r}
set.seed(44)
bootres <- onesampb(g1, est = mean, trim = 0.2, hyp = 0.5, small.n = TRUE, nboot = 1000)
bootres$estimate
bootres$ci
bootres$p.value
```

#### Plot results
```{r}
plot.boot(bootres, label_size = 6)
```

### One-sided
```{r}
set.seed(44)
bootres <- onesampb(g1, est = mean, trim = 0.2, hyp = 0.5, small.n = TRUE, nboot = 1000, alternative = "greater")
bootres$estimate
bootres$ci
bootres$p.value
```

#### Plot results
```{r}
plot.boot(bootres, alternative = "greater", label_height = 0.95, label_size = 5)
```

## Two-sample inference: central tendency 

### Two-sided
```{r}
set.seed(44)
bootres <- twosampb(g1, g2, est = mean, trim = 0.2, hyp = 0, small.n = TRUE, nboot = 1000)
bootres$estimate
bootres$ci
bootres$p.value
```

### Plot results
```{r}
plot.boot(bootres)
```

### Plot bivariate distribution of bootstrap estimates
```{r}
plot.boot2(bootres)
```

### One-sided
```{r}
set.seed(44)
bootres <- twosampb(g1, g2, est = mean, trim = 0.2, hyp = 0, small.n = TRUE, nboot = 1000, alternative = "greater")
bootres$estimate
bootres$ci
bootres$p.value
```

### Plot results
```{r}
plot.boot(bootres, alternative = "greater")
```

## Two-sample inference: spread 

Inference about the interquartile range using `hd.iqr()`.

### Two-sided
```{r}
set.seed(44)
bootres <- twosampb(g1, g2, est = hd.iqr, hyp = 0, small.n = TRUE, nboot = 1000)
bootres$estimate
bootres$ci
bootres$p.value
```

#### Plot results
```{r}
plot.boot(bootres)
```

#### Plot bivariate distribution of bootstrap estimates
```{r}
plot.boot2(bootres)
```

### One-sided
```{r}
set.seed(44)
bootres <- twosampb(g1, g2, est = hd.iqr, hyp = 0, small.n = TRUE, nboot = 1000, alternative = "greater")
bootres$estimate
bootres$ci
bootres$p.value
```

#### Plot results
```{r}
plot.boot(bootres, alternative = "greater")
```




